Flowchart of Current Algorithm

### 2009-05-07 docs/flowchart/flowchart.tex
minor
don't need to add temporal location of g in planar since g is constant
fixed a notational inconsistency
more cleanup
a lot of wordiness in step 8 that doesn't need to be repeated since we say the exact same thing in step 4.  Step 8 now refers to step 4 in certain parts for explanations

\item[plane-parallel:] We discretize the base state density evolution 
equation from paper III equation (29) without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ]
         \nonumber \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are computed using the
procedure described in paper II, Appendix A.\\

@@

\item {\bf Compute $w_0$} is the process by which we compute $w_0$.
The arguments are different for plane-parallel and spherical.
\begin{description}
\item {\bf Compute $w_0$ Planar}$[\Sbar^{\inp},\gammabar^{\inp},p_0^{\inp},\psi^{\inp}] 
\rightarrow [w_0^{\outp}]$:

@@

\item {\bf Compute $w_0$ Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},
\etarho^{\inp}] \rightarrow[w_0^{\outp}]$:

@@

To solve for $\dw$ we will need to substitute for the derivatives of $p_0.$
To do so we start with the equation of hydrostatic equilibrium in a spherical geometry:
\begin{equation}
\frac{\partial p_0}{\partial r} = -\rho_0 g \, ; \quad
g = \frac{G m_\mathrm{encl}}{r^2}\enskip,
\end{equation}
where $m_\mathrm{encl}(r)$ is the mass enclosed at radius $r$ and $G$ is the
gravitational constant.  
Using this, we can then write equation (\ref{eq:sphconstraint_2}) as:
\begin{eqnarray}
\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=&
 \frac{\partial}{\partial t} \left (\rho_0 g \right )
+\frac{\partial}{\partial r} \left (w_0 \rho_0 g \right) \nonumber \\
&=&
  g \left[ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right]
+ \rho_0 \left( \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} \right) \enskip.\nonumber\\
&&
\label{eq:sphconstraint_3}
\end{eqnarray}
The mass enclosed inside any radius, $r$, is $m_\mathrm{encl}(r) = 4 \pi
\int_0^r \rhozero(s) s^2 ds$, or alternately, $\partial m_\mathrm{encl}/\partial r = 4\pi r^2 \rho_0$.  
The Lagrangian derivative of the enclosed mass is then:
\begin{eqnarray}
\frac{D_0 m_\mathrm{encl}}{D t} &=& \frac{\partial m_\mathrm{encl}}{\partial t} + w_0 \frac{\partial m_\mathrm{encl}}{\partial r} \nonumber \\
 &=& 4\pi \left( \frac{\partial}{\partial t} \int_0^r \rhozero(s) s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
&=& 4\pi \left( \int_0^r \frac{\partial \rho_0}{\partial t} s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
&=& 4\pi \left\{ -\int_0^r \left [ \frac{1}{s^2}\frac{\partial(s^2 \rho_0 w_0)}{\partial s} + \frac{1}{s^2}\frac{\partial (s^2 \etarho)}{\partial s} \right ]s^2 ds + w_0 r^2 \rho_0 \right\} \nonumber \\
&=&  \left . 4\pi \left(-s^2 \rho_0 w_0 \right |_0^r - \left . s^2 \etarho \right |_0^r + w_0 r^2 \rho_0 \right) \nonumber \\
&=& -4\pi r^2 \etarho \label{eq:dmdt}
\end{eqnarray}
where we used the spherical form of equation (29) in paper III, 
\begin{equation}
\label{eq:sph_continuity}
\frac{\partial \rhozero}{\partial t} + \frac{1}{r^2} \frac{\partial (r^2 \rho_0 w_0)}{\partial r} 
+ \frac{1}{r^2} \frac{\partial (r^2 \etarho )}{\partial r}  = 0 \enskip ,
\end{equation}
to eliminate $\partial \rho_0/\partial t$.  We note that in the
absence of any mixing, $\etarho=0$, and  $D_0 m_\mathrm{encl}/Dt = 0.$
Equation (\ref{eq:dmdt})
allows us to write the Lagrangian change in the gravitational
acceleration as:
\begin{eqnarray}
\frac{D_0 g}{D t} = \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} &=& \frac{D_0}{Dt}\left(\frac{G m_\mathrm{encl}}{r^2}\right) \nonumber \\
&=& G m_\mathrm{encl}\frac{D_0}{D t}\left(\frac{1}{r^2}\right) + \frac{G}{r^2}\frac{D_0 m_\mathrm{encl}}{Dt} \nonumber \\
&=& -\frac{2 w_0 G m_\mathrm{encl}}{r^3} - 4 \pi G \etarho \nonumber \\
&=& -\frac{2 w_0 g}{r} - 4 \pi G \etarho \enskip .
\end{eqnarray}

@@

\noindent We discretize this elliptic equation in the radial dimension as:
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j} -
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j-1} \right\}
+ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{j-\myhalf} \nonumber\\
&=& 
- \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
\left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
   - \left( \frac{4 \ow \rho_0 g}{r} \right)_{j-\myhalf}
   - \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip , \nonumber
\end{eqnarray}
where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
Then, using hydrostatic equilibrium, we expand this to 
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
 \left( \frac{\gammabar p_0}{r^2}\right)_{j  } \frac{\left[ (r^2 \dw)_{j+\myhalf} - (r^2 \dw)_{j-\myhalf} \right]}{\Delta r}
-\left( \frac{\gammabar p_0}{r^2}\right)_{j-1} \frac{\left[ (r^2 \dw)_{j-\myhalf} - (r^2 \dw)_{j-\thalf}\right]}{\Delta r} \right\} \nonumber \\
&&  - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \dw)_{j-\myhalf}  \nonumber \\
&=&   \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
 - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
  \left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip, \nonumber \\
\end{eqnarray}
where $\etarho$ at cell-centers is the average of the neighboring edge-values.  
If we write this in matrix form, so that:
\begin{equation}
A_j (r^2 \dw)_{j-\thalf} + B_j (r^2 \dw)_{j-\myhalf} + C_j (r^2 \dw)_{j+\myhalf} = F_j\enskip,
\end{equation}
then:
\begin{eqnarray}
A_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1}  \enskip, \nonumber  \\
B_j &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j  }  
               +\left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1} \right] 
              - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) \nonumber \\
C_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j}  \enskip, \nonumber  \\
F_j &=&  \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
 - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
\left[ \left( r^2 \etarho^{\inp} \right)_{j} - \left( r^2 \etarho^{\inp}  \right)_{j-1} \right] 
\nonumber \\
&& - \left(  4 \pi G \rhozero^{\inp} \etarho^{\inp}   \right)_{j-\myhalf} \enskip \nonumber .
\end{eqnarray}
Here,
\begin{equation}
g_{j-\myhalf} = \frac{G}{r_{j-\myhalf}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\myhalf}^3 - r_{k-\thalf}^3) \; \rho_{0,k-1}^{\inp} \enskip.
\end{equation}

We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
corresponds to $j=1,$ by setting:
\begin{eqnarray}
A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
B_1 &=& 1\enskip. \nonumber
\end{eqnarray}
We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $j=N,$
by setting:
\begin{eqnarray}
A_N &=& -1\enskip, \nonumber \\
B_N &=&  1\enskip, \nonumber \\
C_N = F_N &=& 0\enskip. \nonumber
\end{eqnarray}
Finally, $w_0^{\outp} = \ow + \delta w_0$.
\end{description}

@@

If this is the first step of the calculation ($n=0$), we set
\begin{equation}
S^{\myhalf,\star} = \frac{S^0 + S^1}{2} \enskip ,
\end{equation}
where $S^1$ is found through the iterative process that initializes the calculation.
Otherwise, following the method used in our small-scale low Mach number algorithm
(cite BDRWZ), we extrapolate to the half-time using the source
terms at the previous and current time levels
\begin{equation}
S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
\end{equation}
\item Compute
\begin{equation}
\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
\end{equation}
\item Compute $w_0^{n+\myhalf,\star}$:
\begin{description}
\item For plane-parallel,\\
{\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,\psi^{n-\myhalf}] 
\rightarrow [w_0^{n+\myhalf,\star}]$.

If we are using the volume discrepancy term, compute 
$\overline{p^{\inp}} = \overline{p(\rho,h,X_k)^n}$.
\item For spherical,\\
{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,\rhozero^n,p_0^n,
\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
\end{description}

@@

\item Update the species.  Here we do a difference approximation
  to 
  \begin{equation}
  \frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\Ub \rho X_k) =
         \rho {\omegadot}_k \enskip , \label{eq:species}
  \end{equation}
  neglecting the reaction terms.  The update consists of two steps:

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
  for the conservative update of $(\rho X_k)^{(1)}$. 
  Here we predict $\rho^{'(1)} = \rho^{(1)} - \rhozero^n$ and 
  $X_k^{(1)} = (\rho X_k)^{(1)} / \rho^{(1)}$ to time-centered edges to obtain 
  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
  unsplit Godunov procedure, as described in paper II,
  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  We use the base
  state density, $\rho_0$, to convert this to an edge state for $\rho$,
\begin{equation}
\rho^{(1),\nph,\star} = 
\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip .
\label{eq:rhoXedgestate}
\end{equation}
  In plane-parallel, we use 4th-order spatial averaging to compute $\rho_0$ on 
  radial faces.  In spherical, we first map $\rho_0$ to Cartesian cell-centers 
  using the mapping in section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.

@@

\begin{description}
\item For plane-parallel,
\begin{equation}
 \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdotb \er + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
\end{equation}
\item For spherical, first construct 
$\eta_{\rho}^{{\rm cart},\nph,\star} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf,\star}$ using:
\begin{equation}
\eta_{\rho}^{{\rm cart},\nph,\star} = \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \sum_d \left(\uadvone \cdot \eb_d\right)n_d.
\end{equation}
Then, $\etarho^{\nph,\star}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph,\star}$,
\begin{equation}
\etarho^{\nph,\star} = \overline{\eta_{\rho}^{{\rm cart},\nph,\star}}.
\end{equation}
This gives a radial cell-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
radial edges, average the two neighboring radial bins.
\end{description}

@@

\item Compute $\psi$.
\begin{description}
\item For plane-parallel, 
\begin{equation}
\psi_j^{n+\myhalf,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{n+\myhalf,\star} 
+ \eta_{\rho,j+\myhalf}^{n+\myhalf,\star}\right) g^{n+1,\star} \enskip .
\end{equation}
\item For spherical, first compute:
\begin{equation}
\gammabar^{(1)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, 
X_k^{(1)}\right) \right] \enskip ,
\end{equation}
\begin{equation}
\gammabar^{(2),\star} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2),\star}, p_0^{n+1,\star}, 
X_k^{(2),\star}\right) \right] \enskip .
\end{equation}
Then, define a base time and time-centered $\psi$:
\begin{equation}
\psi_j^{n,\star} = 
\left(\gammabar^{(1)} p_0^n\right)_j
\left \{ \Sbar_j^{n+\myhalf,\star} - 
\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip ,
\end{equation}
\begin{eqnarray}
\psi_j^{n+\myhalf,\star} &=& 
\left(\frac{\gammabar^{(1)}+\gammabar^{(2),\star}}{2}\right)_j
\left(\frac{p_0^n+p_0^{n+1,\star}}{2}\right)_j \nonumber \\
&& \left \{ \Sbar_j^{n+\myhalf,\star} - 
\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip .
\end{eqnarray}
\end{description}

@@

\item Compute $w_0^{n+\myhalf}$:
\begin{description}
\item For plane-parallel,\\
{\bf Compute $w_0$ Planar}$[\overline{S^{\nph}},
\gammabar^{n+\myhalf},p_0^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.

>>>>

\item[plane-parallel:] We discretize the base state density evolution 
equation from paper III equation (29) without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ] \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are computed using the
procedure described in paper II, Appendix A.\\

@@

\item {\bf Compute} {\boldmath $w_0$} is the process by which we compute $w_0$.
The arguments are different for plane-parallel and spherical.
\begin{description}
\item {\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\Sbar^{\inp},\gammabar^{\inp},p_0^{\inp},
\psi^{\inp}] \rightarrow [w_0^{\outp}]$:

@@

\item {\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},
\etarho^{\inp}] \rightarrow[w_0^{\outp}]$:

@@

To solve for $\dw$ we will need to substitute for the derivatives of $p_0.$
To do so we start with the equation of hydrostatic equilibrium in a spherical geometry:
\begin{equation}
\frac{\partial p_0}{\partial r} = -\rho_0 g \, ; \quad
g = \frac{G m_\mathrm{encl}}{r^2}\enskip,
\end{equation}
where $m_\mathrm{encl}(r)$ is the mass enclosed at radius $r$ and $G$ is the
gravitational constant.  
Using this, we can then write equation (\ref{eq:sphconstraint_2}) as:
\begin{eqnarray}
\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=&
 \frac{\partial}{\partial t} \left (\rho_0 g \right )
+\frac{\partial}{\partial r} \left (w_0 \rho_0 g \right) \nonumber \\
&=&
  g \left[ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right]
+ \rho_0 \left( \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} \right) \enskip.
\label{eq:sphconstraint_3}
\end{eqnarray}
The mass enclosed inside any radius, $r$, is $m_\mathrm{encl}(r) = 4 \pi
\int_0^r \rhozero(s) s^2 ds$, or alternately, $\partial m_\mathrm{encl}/\partial r = 4\pi r^2 \rho_0$.  
The Lagrangian derivative of the enclosed mass is then:
\begin{eqnarray}
\frac{D_0 m_\mathrm{encl}}{D t} &=& \frac{\partial m_\mathrm{encl}}{\partial t} + w_0 \frac{\partial m_\mathrm{encl}}{\partial r} \nonumber \\
 &=& 4\pi \left( \frac{\partial}{\partial t} \int_0^r \rhozero(s) s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
&=& 4\pi \left( \int_0^r \frac{\partial \rho_0}{\partial t} s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
&=& 4\pi \left\{ -\int_0^r \left [ \frac{1}{s^2}\frac{\partial(s^2 \rho_0 w_0)}{\partial s} + \frac{1}{s^2}\frac{\partial (s^2 \etarho)}{\partial s} \right ]s^2 ds + w_0 r^2 \rho_0 \right\} \nonumber \\
&=&  \left . 4\pi \left(-s^2 \rho_0 w_0 \right |_0^r - \left . s^2 \etarho \right |_0^r + w_0 r^2 \rho_0 \right) \nonumber \\
&=& -4\pi r^2 \etarho \label{eq:dmdt}
\end{eqnarray}
where we used the spherical form of equation (29) in paper III, 
\begin{equation}
\label{eq:sph_continuity}
\frac{\partial \rhozero}{\partial t} + \frac{1}{r^2} \frac{\partial (r^2 \rho_0 w_0)}{\partial r} 
+ \frac{1}{r^2} \frac{\partial (r^2 \etarho )}{\partial r}  = 0 \enskip ,
\end{equation}
to eliminate $\partial \rho_0/\partial t$.  We note that in the
absence of any mixing, $\etarho=0$, and  $D_0 m_\mathrm{encl}/Dt = 0.$
Equation (\ref{eq:dmdt})
allows us to write the Lagrangian change in the gravitational
acceleration as:
\begin{eqnarray}
\frac{D_0 g}{D t} = \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} &=& \frac{D_0}{Dt}\left(\frac{G m_\mathrm{encl}}{r^2}\right) \nonumber \\
&=& G m_\mathrm{encl}\frac{D_0}{D t}\left(\frac{1}{r^2}\right) + \frac{G}{r^2}\frac{D_0 m_\mathrm{encl}}{Dt} \nonumber \\
&=& -\frac{2 w_0 G m_\mathrm{encl}}{r^3} - 4 \pi G \etarho \nonumber \\
&=& -\frac{2 w_0 g}{r} - 4 \pi G \etarho \enskip .
\end{eqnarray}

@@

\noindent We discretize this elliptic equation in the radial dimension as:
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j} -
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j-1} \right\}
+ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{j-\myhalf} \nonumber\\
&=& 
- \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
\left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
   - \left( \frac{4 \ow \rho_0 g}{r} \right)_{j-\myhalf}
   - \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip ,
\end{eqnarray}
where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
Then, using hydrostatic equilibrium, we expand this to 
\begin{eqnarray}
& \frac{1}{\Delta r} \left\{
 \left( \frac{\gammabar p_0}{r^2}\right)_{j  } \frac{\left[ (r^2 \dw)_{j+\myhalf} - (r^2 \dw)_{j-\myhalf} \right]}{\Delta r} -\left( \frac{\gammabar p_0}{r^2}\right)_{j-1} \frac{\left[ (r^2 \dw)_{j-\myhalf} - (r^2 \dw)_{j-\thalf}\right]}{\Delta r} \right\} & \nonumber \\
&  - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \dw)_{j-\myhalf} = \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf} & \nonumber \\
& - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
  \left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip, &
\end{eqnarray}
where $\etarho$ at cell-centers is the average of the neighboring edge-values.  
If we write this in matrix form, so that:
\begin{equation}
A_j (r^2 \dw)_{j-\thalf} + B_j (r^2 \dw)_{j-\myhalf} + C_j (r^2 \dw)_{j+\myhalf} = F_j\enskip,
\end{equation}
then:
\begin{eqnarray}
A_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1}  \enskip, \\
B_j &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j  }  
               +\left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1} \right] 
              - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) \enskip , \\
C_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j}  \enskip, \\
F_j &=&  \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
 - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
\left[ \left( r^2 \etarho^{\inp} \right)_{j} - \left( r^2 \etarho^{\inp}  \right)_{j-1} \right] 
\nonumber \\
&& - \left(  4 \pi G \rhozero^{\inp} \etarho^{\inp}   \right)_{j-\myhalf} \enskip .
\end{eqnarray}
Here,
\begin{equation}
g_{j-\myhalf} = \frac{G}{r_{j-\myhalf}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\myhalf}^3 - r_{k-\thalf}^3) \; \rho_{0,k-1}^{\inp} \enskip.
\end{equation}

We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
corresponds to $j=1,$ by setting:
\begin{eqnarray}
A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
B_1 &=& 1\enskip.
\end{eqnarray}
We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $j=N,$
by setting:
\begin{eqnarray}
A_N &=& -1\enskip, \nonumber \\
B_N &=&  1\enskip, \nonumber \\
C_N = F_N &=& 0\enskip.
\end{eqnarray}
Finally, $w_0^{\outp} = \ow + \delta w_0$.
\end{description}

@@

If this is the first step of the calculation ($n=0$), we set
\begin{equation}
S^{\myhalf,\star} = \frac{S^0 + S^1}{2} \enskip ,
\end{equation}
where $S^1$ is found through the iterative process that initializes the calculation.
Otherwise, following the method used in our small-scale low Mach number algorithm
(cite BDRWZ), we extrapolate to the half-time using the source
terms at the previous and current time levels
\begin{equation}
S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
\end{equation}
\item Compute
\begin{equation}
\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
\end{equation}
\item Compute $w_0^{n+\myhalf,\star}$:
\begin{description}
\item For plane-parallel,\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,
\psi^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.

If we are using the volume discrepancy term, compute 
$\overline{p^{\inp}} = \overline{p(\rho,h,X_k)^n}$.
\item For spherical,\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
\rhozero^n,p_0^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
\end{description}

@@

\item \label{step:species update} 
Update the species.  Here we consider only the advection terms, 
neglecting the reaction terms, and solve a discretized version of
\begin{equation}
\frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\Ub \rho X_k) = 0 \enskip .
\label{eq:species}
\end{equation}
The update consists of two steps:

@@

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
  for the conservative update of $(\rho X_k)^{(1)}$. 
  Here we predict $\rho^{'(1)} = \rho^{(1)} - \rhozero^n$ and 
  $X_k^{(1)} = (\rho X_k)^{(1)} / \rho^{(1)}$ to time-centered edges to obtain 
  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
  unsplit Godunov procedure, as described in paper II,
  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  Convert the 
  perturbational edge state to a full state using
\begin{equation}
\rho^{(1),\nph,\star} = 
\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip .
\label{eq:rhoXedgestate}
\end{equation}
  In plane-parallel, we use 4th-order spatial averaging to compute $\rho_0$ on 
  radial faces.  In spherical, we first map $\rho_0$ to Cartesian cell-centers 
  using the mapping in section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.

@@

\begin{description}
\item For plane-parallel,
\begin{equation}
 \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdotb \er + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
\end{equation}
\item For spherical, first construct 
$\eta_{\rho}^{{\rm cart},\nph,\star} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf,\star}$ using:
\begin{equation}
\eta_{\rho}^{{\rm cart},\nph,\star} = \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \sum_d \left(\uadvone \cdot \eb_d\right)n_d.
\end{equation}
Then, $\etarho^{\nph,\star}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph,\star}$,
\begin{equation}
\etarho^{\nph,\star} = \overline{\eta_{\rho}^{{\rm cart},\nph,\star}}.
\end{equation}
This gives a cell-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
radial edges, average the two neighboring radial cell-centers.
\end{description}

@@

\item Compute $\psi$.
\begin{description}
\item For plane-parallel, 
\begin{equation}
\psi_j^{n+\myhalf,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{n+\myhalf,\star} 
+ \eta_{\rho,j+\myhalf}^{n+\myhalf,\star}\right) g \enskip .
\end{equation}
\item For spherical, first compute:
\begin{equation}
\gammabar^{(1)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, 
X_k^{(1)}\right) \right] \enskip ,
\end{equation}
\begin{equation}
\gammabar^{(2),\star} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2),\star}, p_0^{n+1,\star}, 
X_k^{(2),\star}\right) \right] \enskip .
\end{equation}
Then, define a base time and time-centered $\psi$:
\begin{equation}
\psi_j^{n,\star} = 
\left(\gammabar^{(1)} p_0^n\right)_j
\left \{ \Sbar_j^{n+\myhalf,\star} - 
\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip ,
\end{equation}
\begin{eqnarray}
\psi_j^{n+\myhalf,\star} &=& 
\left(\frac{\gammabar^{(1)}+\gammabar^{(2),\star}}{2}\right)_j
\left(\frac{p_0^n+p_0^{n+1,\star}}{2}\right)_j \nonumber \\
&& \left \{ \Sbar_j^{n+\myhalf,\star} - 
\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip .
\end{eqnarray}
\end{description}

@@

\item Compute $w_0^{n+\myhalf}$:
\begin{description}
\item For plane-parallel,\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph}},
\gammabar^{n+\myhalf},p_0^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.
